marlin-paper-notes

### 2021-09-15 documents/03_marlin_paper.Rmd
safety stop before open access idea

** something to point out: increased catches are a paradox, either come from rebuilding stock or concentrated fishing, but trajectory can tell you something. not possible for MPAs to increase catches without decreasing catches first if they are rebuilding? **

@@

Marine protected areas (MPAs) are a growing part of the marine
conservation toolbox, with 5.9% of the world's oceans currently in implemented MPAs [@grorud-colvert2021], and bodies such as the IUCN calling for 30% protection by
2030. Many of the largest of these MPAs are being implemented in the
form of large MPAs primarily covering pelagic ecosystems. These pelagic systems operate
very differently from the coastal areas where modern MPA design
practices have been developed, characterized by highly mobile species and fishing fleets using a
diverse mix of static and dynamic ocean features. These ecosystems
combinations of relatively abundant species targeted by fishing, such as
most tuna stocks, and highly threatened species caught as incidental
"bycatch" in targeted fishing operations, such as oceanic whitetip
sharks. Blue-water MPAs are increasingly being proposed specifically to
benefit these by-catch populations while minimizing losses or where
possible providing benefits to fisheries.

While the effect of MPAs on fisheries yields is a continuously debated
topic, the conservation impacts of MPAs have received less attention of
late, due in part perhaps to the perceived obvious nature of the
problem: if you prevent fishing of a species within an area, there will
be more of that species in that area, and so the conservation benefits
of MPAs will increase with MPA size. \@hastings2017 clearly illustrates
this idea. Using a simple two-species two-patch analytical model, they
derive conditions under which MPAs can solve the "weak-stock" problem in
which a less productive species is incidentally exploited by a fleet
targeting a more valuable and productive stock.

However, as @hastings2017 point out, their results depend on a very
specific set of assumptions, namely that 1) adults are perfectly
sedentary 2) the target and bycatch species overlap perfectly in terms
of habitat xx....The authors suggest that the impact of these
assumptions should be tested in scenarios under which they are unlikely
to apply. We build on \@hastings2017 to consider the role of MPAs in reducing bycatch reductino while managing tradeoffs with fisheries under a more realistic model reflecting the dynamics of pelagic ecosystems and fisheries.

The core models underpinning MPA science have largely been based on
stylized versions of near-shore reef-style xx what to call that, like
coral / rocky / kelp forest, etc xx systems, where multiple species
often share common habitat, and adults have reasonable site fidelity
while larvae are broadly dispersed. Pelagic ecosystems have vastly
different dynamics than coastal seas [@game2009; @lubchenco2003],
characterized by highly migratory species specialized to take advantage
of highly dynamic ocean features such as eddies currents while also
utilizing static features such as seamounts. Pelagic ecosystems also
support both coastal domestic fishing industries along with highly
mobile international high-seas fishing operations. Given these dynamics,
it remains unclear how many of the lessons from coastal MPA science are
applicable to pelagic ecosystems [\@ovando2021XX].

Pelagic ecosystems present both opportunities and challenges for MPAs.
The highly productive and mobile nature of targeted fisheries in these
systems may mean that highly targeted MPAs can be designed that provide
effective conservation at minimal fishery costs. Strategic protection of
key aggregation locations may enhance depleted pelagic fisheries [xx
that Indian ocean study I can't remember right now]. xx big question
here: how focused do we want this paper to be? is it just about
conservation tradeoffs or lack thereof with bycatch species, or is about
the net conservation and fishery effects of BWMPAs. I'm inclined the
former to tee-up the later xx.

However, the dynamic nature of pelagic systems also creates a risk of unintended conservation consequences of MPAs. When an MPA is implemented on former fishing grounds, fishing fleets have two choices: exit the fishery or move to new fishing grounds. When MPAs redistribute effort away from one area and into another, they can accidentally increase fishing effort on species found within the new fishing grounds [@abbott2012]. This problem of policy "leakage" is well known in many environmental fields [@citations] but the general properties of this problem have received relatively little attention in the theoretical or empirical MPA literature [see @abbott2012;@savina2008;fulton2011a;@baum2003 for notable exceptions]. We fill this gap by examining the general range of conditions...

- Assessing the conditions under which this may be a problem using a generalized simulation environment
  - General insights about design strategy, effect of varying degrees of habitat overlap, etc, tradeoffs with fisheries

- Given that "it depends" nature of the problem, assess the specific potential for this problem in a case study
  - specific insights about design strategy, effect of varying degrees of habitat overlap, etc, tradeoffs with fisheries

Our goal here is not to provide tactical recommendations on the design
of blue-water MPAs, but rather to provide a strategic evaluation of the
sensitivity of bycatch populations and targeted fisheries to key
unknowns and design choices in MPA planning.

@@

@hastings2017 employed a simple model to derive conditions under which
MPAs would benefit bycatch species. However they also pointed out that
even small increases in model complexity quickly make analytical
solutions to the effects of MPAs intractable. As such, we based our numerical
analysis on a spatially explicit, multi-species, multi-fleet,
bio-economic model we call
[`marlin`](https://github.com/DanOvando/marlin). Models of MPAs span a
wide range of complexity, from simple analytical tools to end-to-end
ecosystem models [@fulton2015]. `marlin` follows a "models of
intermediate complexity" (MICE) philosophy [@plagányi2014] targeted
specifically around the role of technical interactions across species.
Many MPA models simulate the impacts of MPAs on a single-species level [@2021]. This can be useful for for example examining the ways in which species-specific life history traits interact with fleet dynamics to influence MPA outcomes. However, bycatch problems are driven by technical interactions, in which fishing effort for a target species incidentally captures another species xx how do we want to talk about this in here? Like usually incidentally, but clearly sometimes people try and catch sharks too xx. `marlin` accounts for this by allowing the user to simulate an arbitrary number of fleets capable of catching an arbitrary number of species in a spatial area. Different fleets can place different economic value on different species, while also having species-specific selectivity functions. This allows us to simulate bycatch dynamics of MPAs in pelagic ecosystems.

We first ran a series of simulation exeriments designed to explore broad sensitivities of bycatch and fisheries to key factors.

We then ran a focused case study intended to reflect some of the traits of the tuna fisheries of the WCPO, holding constant...

@@

- Design of movement and habitat
  - Movement is a product of distance and habitat, such that critters will all else equal move to the closest patches with the best habitat

@@

-

## Simulation Experiments

Given the complexity of the operating model used here, a complete sensitivity analysis across all parameters is not practical. Instead, we focus on the role of life history, fishing pressure,

Fleets exert a given amount of effort in a patch each time step. The catch resulting from that effort is given by the Baranov [@baranov18XX] equatino

@@

@@

In order to constrain the parameter space somewhat, we also ran a series of case study experiments. Rathere than interating over different habitat configurations and movement dynamics, we fix habitat for each species based on the results of @bradley2021XX, and set ontogenetic movement and seasonal migrations to FALSE.

@@

Each model run started at equilibrium conditions and was then run for `r years` years (`r seasons * years` seasons. We assessed performance by comparing relevant metrics (e.g. depletion and revenue) under a given scenario with MPAs relative to the values of those same metrics in the simulated counterfactual world identical in every way save for the presencee of MPAs.

@@

## Simulation Experiments

@@

However, three important features do emerge. The first is that life history plays an important role in the potential for unintended conservation consequences from MPAs. All else equal, MPAs were much less likely to produce losses in spawning biomass for our tuna-like species than the marlin or shark-like species. Given that within the broad life history traits we do vary both adult and larval movement, as well as the timing of density dependence, we believe this result is mostly due to the high resilience (high steepness, low age at maturity, fast growth) of the tuna-like species relative to the marlin and shark-like species.

@@

The third is that extreme outcomes (both positive and negative), particularly at lower MPA sizes, are largely driven by the presence of small range sizes interacting with seasonal migrations. Across species simulations without seasonal migrations and small ranges were much more likely to produce extreme results than those with seasonal migrations and/or larger ranges. Presence of seasonal migrations acts as a dilutant: As a given static MPA only provides protection, or displaces fishing effort, onto a particular area in only half of the year, both the benefits and costs of MPAs are smaller. When seasonal migrations are not present and the species range is small, MPAs are more capable of providing either large benefits or large losses.

@@

@@

- Under the levels of fishing mortality in the WCPO, optimization and design can help reduce tradeoffs between fisheries revenue and unintended consequences than others, but macro-level tradeoffs between biodiversity and fishery revenue are very insensitive to design choices, with a pretty clear "increasing biodiversity through MPAs reduces fishery revenues". 
  -  xx This is clearly hyper-sensitive to the one thing we don't iterate over much here: how overfished are the target species. At the moment "they're not". But, this runs fast enough, is it worth running a hypothetical case study where they are just to illustrate how things change?

@@

- Should pay particular attention to potential though when species near tipping points are of concern: hard to imagine those not being part of the prioritization but something to consider. 
- Other interesting things.

@@

Thinking through the “naive” MPA design strategies to compare against the optimized outcomes. Thoughts at the moment below. These can be run in a few seconds each, so we cab be be pretty liberal in coming up with combinations if there are other ideas people want to try, but also if some of these seem pointless can drop as well!
6:11
Design strategies
1. Assign protection by SSB / SSB0 weighted abundance in space (protect the highest concentrations of the most threatened species)
2. Assign protection by bycatch rates (distribution of biomass bycatch / biomass target)
3. Assign protection by targeting fishing effort
4. Assign protection by avoiding fishing effort
Knowledge states
1. Degrees of accuracy over the thing being measured (i.e. error in the design strategy)
2. Number of species considered (e.g. design aronnd only 2 bycatch species instead of all)
3. über naive: just close off random patches up to a specified size
We then plot these points in objective function space and compare to the efficiency frontier / consider frequency and magnitude of conservation and economic tradeoofs
The question all this is answering: how sensitive are conservation and economic outcomes to design strategy / uncertainty when there are technical interactions and heterogeneity in habitat? The first stage is teeing up the idea that beyond conservation / yield tradeoffs, as you increase habitat heterogeneity with technical interactions you increase the potential for conservation tradeoffs. But, the degree is going to be super context dependent, so to dive in a bit we then focus on this case study of the particular species / habitat overlap in the WCPO. Further work might reveal more generalities in this, but htis is basically getting a sense for the potential magnitude of the challenge.

@@

"
Spatial management was insufficient as the dominant management lever because it proves quite difficult to strike the balance between protecting areas that are of a size sufficient to produce any conservation benefit and providing sufficient fishing grounds to sustain an econom- ically viable industry (which is not possible in many fisheries if closures on the order of 70–80% or more are imposed, Fulton et al. 2007). More- over, Savina etal. (2008) found that effort displaced by closures potentially undermines any benefits coming from the zoning. For instance, strong improvements in diversity (by nearly 20%) under heavy fishing pressure occurred if effort outside a marine protected area (MPA) was controlled. However, without effort control, there was a significant (>20%) drop in overall diversity. Fulton et al. (2007), Savina et al. (2008), Smith et al. (2010) and unpublished work associated with Horne et al. (2010) all found that while spatial management appears to be an effective means of satisfying conservation objectives con- cerning habitat and restoring age-structures of less mobile species, benefits from closures for individual species depend on its trophic position, degree of mobility, geographical distribution and extent of ontogenetic shifts in habitat use and depth (highly mobile or prey species do not benefit in the way that more site attached or predatory species do)."

>>>>

\*\* something to point out: increased catches are a paradox, either
come from rebuilding stock or concentrated fishing, but trajectory can
tell you something. not possible for MPAs to increase catches without
decreasing catches first if they are rebuilding? \*\*

@@

Marine protected areas (MPAs) are a growing part of the marine
conservation toolbox, with 5.9% of the world's oceans currently in
implemented MPAs [@grorud-colvert2021], and bodies such as the IUCN
calling for 30% protection by 2030. The modern science of MPAs was
largely developed around coastal ecosystems, often characterized by
relatively static features such as coral reefs [citation]. However,
large MPAs are increasingly being implemented and considered across
pelagic ecosystems. These pelagic systems are characterized by highly
mobile species and fishing fleets using a diverse mix of static and
dynamic ocean features [@game2009; @lubchenco2003]. These ecosystems are
also home to combinations of relatively abundant species targeted by
fishing, such as many tuna stocks [@pons2016], and highly threatened
species caught as incidental "bycatch" in targeted fishing operations,
such as oceanic whitetip sharks [@pacoureau2021]. Large pelagic MPAs are
increasingly being proposed specifically to benefit these by-catch
populations while minimizing losses or where possible providing benefits
to fisheries [citation].

In this paper we evaluate the potential conservation and fishery effects
of MPAs within the pelagic context. We focus in particular on the
potential for pelagic MPAs to provide conservation benefits to bycatch
species. While the effect of MPAs on fisheries yields is a continuously
debated topic, the conservation impacts of MPAs are seemingly simpler,
albeit challenging to detect in many cases [@ovando2021a], due in part
perhaps to the perceived obvious nature of their benefits: if you
prevent fishing of a species within an area, there will be more of that
species in that area, and so the conservation benefits of MPAs will
increase with MPA size. \@hastings2017 clearly illustrates this idea.
Using a simple two-species two-patch analytical model, they derive
conditions under which MPAs can solve the "weak-stock" problem in which
a less productive species is caught as btcatch by a fleet targeting a
more valuable and productive stock.

However, as @hastings2017 point out, their results depend on a very
specific set of assumptions, namely that 1) adults are perfectly
sedentary 2) the target and bycatch species overlap perfectly in terms
of habitat and 3) recruits are perfectly mixed throughout the system.
Pelagic species are likely to violate one or more fo these assumptions.
We build on \@hastings2017 to consider the role of MPAs in reducing
bycatch while managing tradeoffs with fisheries under a more resolved
model reflecting the dynamics of pelagic ecosystems and fisheries.

Pelagic ecosystems present both opportunities and challenges for MPAs.
The highly productive and mobile nature of targeted fisheries in these
systems may mean that highly targeted MPAs can be designed that provide
effective conservation at minimal fishery costs. Strategic protection of
key aggregation locations may enhance depleted pelagic fisheries [xx
that Indian ocean study I can't remember right now]. However, the
dynamic nature of pelagic systems also creates a risk of unintended
conservation consequences of MPAs. When an MPA is implemented on former
fishing grounds, fishing fleets have two choices: exit the fishery or
move to new fishing grounds. When MPAs redistribute effort away from one
area and into another, they can accidentally increase fishing effort on
species found within the new fishing grounds [@abbott2012]. This problem
of policy "leakage" is well known in many environmental fields
[@citations] but the general properties of this problem have received
relatively little attention in the theoretical or empirical MPA
literature [see
@abbott2012;@savina2008;[fulton2011a;\@baum2003](mailto:fulton2011a;@baum2003){.email}
for notable exceptions].

This study addresses this gap using a novel bio-economic simulation
model, capable of reflecting several of the key challenges of pelagic
ecosystems, namely high and variable movement rates, mobile fishing
fleets, and heterogeneous species distributions. We first provide a set
of stylized simulations intended to illustrate the potential and
challenges of achieving conservation objectives through pelagic MPAs,
and highlight particularly important drivers and the magnitude and
direction of MPA effects. We then present a more focused case study
based on the Western and Central Pacific Ocean, examining the potential
outcomes of pelagic MPAs on bycatch and target species.

<!-- We fill this gap by examining the general range of conditions… -->

\<!-- \# Assessing the conditions under which this may be a problem
using a generalized simulation environment

General insights about design strategy, effect of varying degrees of
habitat overlap, etc, tradeoffs with fisheries

-   Given that "it depends" nature of the problem, assess the specific
    potential for this problem in a case study

    -   specific insights about design strategy, effect of varying
        degrees of habitat overlap, etc, tradeoffs with fisheries

Our goal here is not to provide tactical recommendations on the design
of blue-water MPAs, but rather to provide a strategic evaluation of the
sensitivity of bycatch populations and targeted fisheries to key
unknowns and design choices in MPA planning. --\>

@@

@hastings2017 employed an analytically tractable model to derive
conditions under which MPAs would benefit bycatch species. However they
also pointed out that even small increases in model complexity quickly
make analytical solutions to the effects of MPAs intractable. As such,
we employ a numerical simulation strategy in the form of a spatially
explicit, multi-species, multi-fleet, bio-economic model we call
[`marlin`](https://github.com/DanOvando/marlin). Models of MPAs span a
wide range of complexity, from simple analytical tools to end-to-end
ecosystem models [@fulton2015]. `marlin` follows a "models of
intermediate complexity" (MICE) philosophy [@plagányi2014] targeted
specifically around the role of technical interactions across species.
Many MPA models simulate the impacts of MPAs on a single-species level
[@ovando2021a]. This can be useful for for example examining the ways in
which species-specific life history traits interact with fleet dynamics
to influence MPA outcomes. However, bycatch problems are driven by
technical interactions, in which fishing effort for a target species
incidentally captures another species. `marlin` accounts for this by
allowing the user to simulate an arbitrary number of fleets capable of
catching an arbitrary number of species in a spatial area. Different
fleets can place different economic value on different species, while
also having species-specific selectivity functions. This allows us to
simulate bycatch dynamics of MPAs in pelagic ecosystems.

@@

-   Design of movement and habitat

    -   Movement is a product of distance and habitat, such that
        critters will all else equal move to the closest patches with
        the best habitat

@@

Fleets exert a given amount of effort in a patch each time step. The
catch resulting from that effort is given by the Baranov [@baranov1918]
equation

@@

###

@@

## Illustrative Experiments

Given the complexity of the operating model used here, a complete
sensitivity analysis across all parameters is not practical. Instead, we
focus on the role of life history, fishing pressure, and MPA design.

## WCPO Case Study

In order to constrain the parameter space somewhat, we also ran a series
of case study experiments. Rather than iterating over different habitat
configurations and movement dynamics, we defined habitat based on
reported catch per unit effort (CPUE) from the WCPO. XX more methods
here XX

@@

Each model run started at equilibrium conditions and was then run for
`r years` years (`r seasons * years` seasons. We assessed performance by
comparing relevant metrics (e.g. depletion and revenue) under a given
scenario with MPAs relative to the values of those same metrics in the
simulated counterfactual world identical in every way save for the
presence of MPAs.

@@

## Illustrative Experiments

@@

However, three important features do emerge. The first is that life
history plays an important role in the potential for unintended
conservation consequences from MPAs. All else equal, MPAs were much less
likely to produce losses in spawning biomass for our tuna-like species
than the marlin or shark-like species. Given that within the broad life
history traits we do vary adult and larval movement, fishing mortality,
as well as the timing of density dependence, we believe this result is
mostly due to the high resilience (high steepness, low age at maturity,
fast growth) of the tuna-like species relative to the marlin and
shark-like species.

@@

The third is that extreme outcomes (both positive and negative),
particularly at lower MPA sizes, are largely driven by the presence of
small range sizes interacting with seasonal migrations. Across species,
simulations without seasonal migrations and small ranges were much more
likely to produce extreme results than those with seasonal migrations
and/or larger ranges. Presence of seasonal migrations acts as a
dilutant: As a given static MPA only provides protection, or displaces
fishing effort, onto a particular area in only half of the year, both
the benefits and costs of MPAs are smaller. When seasonal migrations are
not present and the species range is small, MPAs are more capable of
providing either large benefits or large losses.

@@

motivated by but not limited to pelagic ecosystems

@@

-   Under the levels of fishing mortality in the WCPO, optimization and
    design can help reduce tradeoffs between fisheries revenue and
    unintended consequences than others, but macro-level tradeoffs
    between biodiversity and fishery revenue are very insensitive to
    design choices, with a pretty clear "increasing biodiversity through
    MPAs reduces fishery revenues".

    -   xx This is clearly hyper-sensitive to the one thing we don't
        iterate over much here: how overfished are the target species.
        At the moment "they're not". But, this runs fast enough, is it
        worth running a hypothetical case study where they are just to
        illustrate how things change?

@@

-   Should pay particular attention to potential though when species
    near tipping points are of concern: hard to imagine those not being
    part of the prioritization but something to consider.

-   Other interesting things.

@@

Thinking through the "naive" MPA design strategies to compare against
the optimized outcomes. Thoughts at the moment below. These can be run
in a few seconds each, so we cab be be pretty liberal in coming up with
combinations if there are other ideas people want to try, but also if
some of these seem pointless can drop as well! 6:11 Design strategies 1.
Assign protection by SSB / SSB0 weighted abundance in space (protect the
highest concentrations of the most threatened species) 2. Assign
protection by bycatch rates (distribution of biomass bycatch / biomass
target) 3. Assign protection by targeting fishing effort 4. Assign
protection by avoiding fishing effort Knowledge states 1. Degrees of
accuracy over the thing being measured (i.e. error in the design
strategy) 2. Number of species considered (e.g. design aronnd only 2
bycatch species instead of all) 3. über naive: just close off random
patches up to a specified size We then plot these points in objective
function space and compare to the efficiency frontier / consider
frequency and magnitude of conservation and economic tradeoofs The
question all this is answering: how sensitive are conservation and
economic outcomes to design strategy / uncertainty when there are
technical interactions and heterogeneity in habitat? The first stage is
teeing up the idea that beyond conservation / yield tradeoffs, as you
increase habitat heterogeneity with technical interactions you increase
the potential for conservation tradeoffs. But, the degree is going to be
super context dependent, so to dive in a bit we then focus on this case
study of the particular species / habitat overlap in the WCPO. Further
work might reveal more generalities in this, but htis is basically
getting a sense for the potential magnitude of the challenge.

@@

" Spatial management was insufficient as the dominant management lever
because it proves quite difficult to strike the balance between
protecting areas that are of a size sufficient to produce any
conservation benefit and providing sufficient fishing grounds to sustain
an econom- ically viable industry (which is not possible in many
fisheries if closures on the order of 70--80% or more are imposed,
Fulton et al. 2007). More- over, Savina etal. (2008) found that effort
displaced by closures potentially undermines any benefits coming from
the zoning. For instance, strong improvements in diversity (by nearly
20%) under heavy fishing pressure occurred if effort outside a marine
protected area (MPA) was controlled. However, without effort control,
there was a significant (\>20%) drop in overall diversity. Fulton et al.
(2007), Savina et al. (2008), Smith et al. (2010) and unpublished work
associated with Horne et al. (2010) all found that while spatial
management appears to be an effective means of satisfying conservation
objectives con- cerning habitat and restoring age-structures of less
mobile species, benefits from closures for individual species depend on
its trophic position, degree of mobility, geographical distribution and
extent of ontogenetic shifts in habitat use and depth (highly mobile or
prey species do not benefit in the way that more site attached or
predatory species do)."

### 2021-09-15 documents/marlin-paper-notes.Rmd
safety stop before open access idea

>>>>

# comparing catches to CPUE to abundance

```{r}

rfmo <- "wcpfc"
mats <- list.files(path = here("data", "tuna-catch"))

# mats <- mats[str_detect(mats, "wcpfc_")]

mats <- mats[str_detect(mats, paste0(rfmo,"_")) & str_detect(mats, "tonnes") & str_detect(mats, "monthly") & str_detect(mats, "2017") & str_detect(mats, "5grid")]

get_layer <- function(file) {
  # file <- mats[7]

  file_comps <- str_split(file, "_", simplify = TRUE)

  species <- paste(file_comps[,7],file_comps[,8])

  species <- str_replace(species, ".tif", "")
# 
  hab <- stars::read_stars(here("data", "tuna-catch", file)) %>%
    sf::st_as_sf() %>%
    recenter_vector(center= -200)

#   plot(hab)

  centers <- sf::st_centroid(hab)

  centers <- sf::st_coordinates(centers) %>% 
    as.data.frame()

  centers$catch = as.data.frame(hab)[,1]

  x_binsize <- (max(centers$X) - min(centers$X) + 1) / resolution

  y_binsize <- (max(centers$Y) - min(centers$Y) + 1) / resolution

  if (x_binsize < 1 | y_binsize < 1) {
    stop("resolution is too high for input data")
  }

  habitat <- centers %>%
    mutate(X = X - min(X),
           Y = Y - min(Y)) %>%
    mutate(rough_x = floor(X / x_binsize),
           rough_y = floor(Y / y_binsize)) %>%
    dplyr::select(contains("rough_"), catch)

  habitat <- habitat %>%
    group_by(rough_x, rough_y) %>%
    summarise(hab = sum(catch)) %>%
    mutate(xy = rough_x * rough_y)

  mod <-
    gamm4::gamm4(hab ~ s(rough_x) + s(rough_y) + s(xy), data = habitat)

  habitat_frame <-
    expand_grid(rough_x = 1:resolution,
                rough_y = 1:resolution) %>%
    mutate(xy = rough_x * rough_y)

interp_habitat <- predict(mod$gam, newdata = habitat_frame)

interp_habitat <- pmax(0, interp_habitat / max(interp_habitat))

habitat_frame$habitat <- interp_habitat

habitat_frame %>%
  ggplot(aes(rough_x, rough_y, fill = habitat)) +
  geom_tile()

  hab <- habitat_frame %>%
    dplyr::select(contains("rough"), habitat) %>%
    rename(x = rough_x, y = rough_y) %>%
    ungroup()

  out <- tibble(scientific_name = species, habitat = list(hab))

}

mats <- map(mats, safely(get_layer))

mats_worked <- map(mats, "error") %>% map_lgl(is.null)

mats <- mats[mats_worked]

mats <- map(mats, "result")

catch_mats <- mats %>%
  bind_rows() %>% 
  unnest(cols = habitat) %>% 
  group_by(scientific_name, x, y) %>% 
  summarise(habitat = sum(habitat)) %>% 
  ungroup() %>% 
  nest(habitat = c(-scientific_name)) %>% 
  ungroup()

catch_habitat <- catch_mats %>%
  bind_rows() %>%
  unnest(cols = habitat) %>%
  ggplot(aes(x, y, fill = habitat)) +
  geom_tile() +
  facet_wrap( ~ scientific_name) +
  scale_fill_viridis_c()

catch_habitat

```

Compare to the SDM stuff

```{r}

mats <- list.files(path = here("data", "matrices"))

mats <- mats[str_detect(mats, "longline_wcpfc")]

get_layer <- function(file) {
  # file <- mats[3]

  file_comps <- str_split(file, "_", simplify = TRUE)

  species <- str_replace(file_comps[, 1], "-", " ")

  fleet <- file_comps[, 2]

  hab <-
    read_csv(here("data", "matrices", file),
             col_names = FALSE,
             skip = 1) %>%
    rename(lat = X1) %>%
    mutate(lat = 1:nrow(.)) %>%
    pivot_longer(
      -lat,
      names_to = "lon",
      values_to = "hab",
      names_prefix = "X"
    ) %>%
    mutate(hab = tidyr::replace_na(hab, 0),
           lon = as.numeric(lon) - 1)

  # hab %>%
  #   ggplot(aes(lon, lat, fill = hab)) +
  #   geom_tile()
  #

  x_binsize <- (max(hab$lon) - min(hab$lon) + 1) / resolution

  y_binsize <- (max(hab$lat) - min(hab$lat) + 1) / resolution

  if (x_binsize < 1 | y_binsize < 1) {
    stop("resolution is too high for input data")
  }

  hab <- hab %>%
    mutate(X = lon - min(lon),
           Y = lat - min(lat)) %>%
    mutate(rough_x = floor(X / x_binsize),
           rough_y = floor(Y / y_binsize)) %>%
    select(contains("rough_"), hab)
  #
  #   hab %>%
  #     ggplot(aes(rough_x, rough_y, fill = hab)) +
  #     geom_tile()

  hab <- hab %>%
    group_by(rough_x, rough_y) %>%
    summarise(hab = sum(hab)) %>%
    mutate(xy = rough_x * rough_y)

  mod <-
    gamm4::gamm4(hab ~ s(rough_x) + s(rough_y) + s(xy), data = hab)

  hab$interp_cpue <- as.numeric(predict(mod$gam))

  hab$habitat <- pmax(0, hab$interp_cpue / max(hab$interp_cpue))

  if (sqrt(nrow(hab)) != resolution) {
    stop("habitat layer does not match simulation resolution")
  }
  hab <- hab %>%
    select(contains("rough"), habitat) %>%
    rename(x = rough_x, y = rough_y) %>%
    ungroup()

  # hab %>%
  #   ggplot(aes(rough_x, rough_y, fill = habitat)) +
  #   geom_tile()

  out <- tibble(scientific_name = species, habitat = list(hab))

}

mats <- map(mats, get_layer)

sdm_habitat <- mats %>%
  bind_rows() %>%
  unnest(cols = habitat) %>%
  filter(scientific_name %in% unique(catch_mats$scientific_name)) %>% 
  ggplot(aes(x, y, fill = habitat)) +
  geom_tile() +
  facet_wrap( ~ scientific_name) +
  scale_fill_viridis_c()

sdm_habitat

catch_habitat / sdm_habitat
```
